A Decision Procedure for (Co)datatypes in SMT Solvers

### 2015-02-17 tex/body.tex
more

%\paragraph{Related Work.}
Although it was written several years ago,
Barrett et al.\ \cite{barrett-et-al-2007} still provides a good account of
related work. Since then, datatypes have been added to the SMT solver
Z3 in unpublished work by Leonardo de Moura and to a SPASS-like prototypical
superposition prover called Pirate by Daniel Wand.
Closely related is the automatic structural
induction in CVC4 by Reynolds and Kuncak \cite{reynolds-kuncak-2015} and in
the SPASS-like prototype by Wand and Weidenbach \cite{wand-weidenbach-201x};
both methods naturally depend on a notion of datatype.
% \textbf{TODO:} http://frocos2013.loria.fr/slides/Peltier.pdf
% \textbf{TODO:} Others (or at least add "e.g.")

@@

If $t$ is a selector term of the form $\const s^k_{\negvthinspace j}( u )$,
since \rn{Split} does not apply to $\Fc$,
$\ec{u}$ must contain a term of the form $\const C_{\negvthinspace\jPrime}\bigl( \const s^1_{\negvthinspace\jPrime}( u ), \ldots, \const s^{n}_{\negvthinspace\jPrime}( u ) \bigr)$ for some $\jPrime$.
%By Lemma~\ref{lem:model-completion}\PART{2} and due to our construction of $\ValC$, we have $\ValC \ec{u}$ is of the form
%$\MU {\Varec{u}}.\; \const C_{\negvthinspace\jPrime}( w_1, \ldots, w_n )$,
%where $\interp{\ValC \ec{u} }{ w_i } = \ValC \ec{ \const s^i_{\negvthinspace\jPrime}( u )}$ for $i \in [n]$.
Since \rn{Inject} and \rn{Clash} do not apply to $\Fc$,
by construction $\ValC$ must be of the form
$\MU {\Varec{u}}.\; \const C_{\negvthinspace\jPrime}( w_1, \ldots, w_n )$,
where for each $i \in [n]$,
$\muvar(w_i) = \smash{\Varec{\vphantom{I}\smash{\const s^i_{\negvthinspace\jPrime}( u )}}}$ and thus by Lemma~\ref{lem:model-completion}\PART{2},
$\interp{\ValC \ec{u}}{ w_i } = \ValC \ec{ \smash{\const s^i_{\negvthinspace\jPrime}( u )}}$.
If $j = \jPrime$, then $\M( t )$ is $\alpha$-equivalent with $\interp{\ValC \ec{u}}{w_k}$, which is equal to $\ValC \ec{ \smash{ \const s^k_{\negvthinspace j}( u )}}$,
which is $\ValC \ec{t}$.
If $j \neq \jPrime$, since \rn{Cong} does not apply,
$\smash{\const s^k_{\negvthinspace j}( u )}$ and $\smash{\const s^k_{\negvthinspace j}( u' )}$
are in distinct equivalence classes only if $u$ and $u'$ are in distinct equivalence classes.
Due to the induction hypothesis and Lemma~\ref{lem:model-completion}\PART{3}, $\M( u ) \neq \M( u' )$ for all such $u$, $u'$.
Thus, we may interpret $\smash{\M(\const s^k_{\negvthinspace j})( \M( u ) )}$ as the value in $\M( \tau )$ that is $\alpha$-equivalent with $\ValC \ec{ t }$.

@@

\rem{TODO: * sharing is caring (let's cite it for a change)}
\rem{TODO: ``theory combination more complicated (and Theorem Completeness too weak)''?}

>>>>

\paragraph{Related Work.}
Although it was written several years ago,
Barrett et al.\ \cite{barrett-et-al-2007} still provides a good account of
related work. Since then, datatypes have been added to the SMT solver
Z3 in unpublished work by Leonardo de Moura and to a SPASS-like prototypical
superposition prover called Pirate by Daniel Wand.
Closely related is the automatic structural
induction in CVC4 by Reynolds and Kuncak \cite{reynolds-kuncak-2015} and in
the SPASS-like prototype by Wand and Weidenbach \cite{wand-weidenbach-201x};
both methods naturally depend on a notion of datatype.
% \textbf{TODO:} http://frocos2013.loria.fr/slides/Peltier.pdf
% \textbf{TODO:} Others (or at least add "e.g.")
% \textbf{TODO:} Nitpick

@@

If $t$ is a selector term of the form $\const s^k_{\negvthinspace j}( u )$,
since \rn{Split} does not apply to $\Fc$,
$\ec{u}$ must contain a term of the form $\smash{\const C_{\negvthinspace\jPrime}\bigl( \const s^1_{\negvthinspace\jPrime}( u ), \ldots, \const s^{n}_{\negvthinspace\jPrime}( u ) \bigr)}$ for some $\jPrime$.
%By Lemma~\ref{lem:model-completion}\PART{2} and due to our construction of $\ValC$, we have $\ValC \ec{u}$ is of the form
%$\MU {\Varec{u}}.\; \const C_{\negvthinspace\jPrime}( w_1, \ldots, w_n )$,
%where $\interp{\ValC \ec{u} }{ w_i } = \ValC \ec{ \const s^i_{\negvthinspace\jPrime}( u )}$ for $i \in [n]$.
Since \rn{Inject} and \rn{Clash} do not apply to $\Fc$,
by construction $\ValC$ must be of the form
$\MU {\Varec{u}}.\; \const C_{\negvthinspace\jPrime}( w_1, \ldots, w_n )$,
where for each $i \in [n]$,
$\muvar(w_i) = \smash{\Varec{\vphantom{I}\smash{\const s^i_{\negvthinspace\jPrime}( u )}}}$ and thus by Lemma~\ref{lem:model-completion}\PART{2},
$\interp{\ValC \ec{u}}{ w_i } = \ValC \ec{ \smash{\const s^i_{\negvthinspace\jPrime}( u )}}$.
If $j = \jPrime$, then $\M( t )$ is $\alpha$-equivalent with $\interp{\ValC \ec{u}}{w_k}$, which is equal to $\ValC \ec{ \smash{ \const s^k_{\negvthinspace j}( u )}}$,
which is $\ValC \ec{t}$.
If $j \neq \jPrime$, since \rn{Cong} does not apply,
$\smash{\const s^k_{\negvthinspace j}( u )}$ and $\smash{\const s^k_{\negvthinspace j}( u' )}$
are in distinct equivalence classes only if $u$ and $u'$ are in distinct equivalence classes.
Due to the induction hypothesis and Lemma~\ref{lem:model-completion}\PART{3}, $\M( u ) \neq \M( u' )$ for all such $u$, $u'$.
Thus, we may interpret $\smash{\M(\const s^k_{\negvthinspace j})( \M( u ) )}$ as the value in $\M( \tau )$ that is $\alpha$-equivalent with $\ValC \ec{ t }$.

@@

\rem{TODO: * sharing is caring (let's cite it for a change)}
\rem{TODO: ``theory combination more complicated (and Theorem Completeness too weak)''?}
\rem{TODO: representation of $\mu$-terms?}
